---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1 - Load the data

```{r}


if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityDataCsv <- read.csv('activity.csv')
str(activityDataCsv)

```

2 - Process/transform the data (if necessary) into a format suitable for your analysis


## What is mean total number of steps taken per day?

```{r}
stepsByDay <- tapply(activityDataCsv$steps, activityDataCsv$date, sum, na.rm=TRUE)
summary(stepsByDay)


```

1 - Make a histogram of the total number of steps taken each day

```{r}
library(ggplot2)


qplot(stepsByDay, xlab='Steps taken each day', ylab='Frequency')

```

2 - Calculate and report the mean and median total number of steps taken per day

```{r}
meanStepsByDay <- mean(stepsByDay)
medianStepsByDay <- median(stepsByDay)
print(meanStepsByDay)
print(medianStepsByDay)
```
## What is the average daily activity pattern?

```{r}
averageStepsPerTimeChunk <- aggregate(x=list(meanSteps=activityDataCsv$steps), by=list(interval=activityDataCsv$interval), FUN=mean, na.rm=TRUE)

str(averageStepsPerTimeChunk)

```

1 - Make a time series plot of the 5-minute interval (x-axis) and the average number
of steps taken, averaged across all days (y-axis)


```{r}
ggplot(data=averageStepsPerTimeChunk, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-min interval") +
    ylab("Average number of steps") 

```

2 - Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps ?

```{r}
maxNumberOfSteps <- which.max(averageStepsPerTimeChunk$meanSteps)
timeMaxNumberOfSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeChunk[maxNumberOfSteps,'interval'])

print(maxNumberOfSteps)

print(timeMaxNumberOfSteps)

```


## Imputing missing values

1 - Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)


```{r}
numberOfMissingValues <- length(which(is.na(activityDataCsv$steps)))

print(numberOfMissingValues)

```

2 - Devise a strategy for filling in all of the missing values in the dataset.

3 - Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}
library(Hmisc)

activityDataCsvImputed <- activityDataCsv
activityDataCsvImputed$steps <- impute(activityDataCsv$steps, fun=mean)
str(activityDataCsvImputed)

```

4a - Make a histogram of the total number of steps taken each day and ...

```{r}
totalStepsByDayImputed <- tapply(activityDataCsvImputed$steps, activityDataCsvImputed$date, sum)


qplot(totalStepsByDayImputed, xlab='Total steps taken each day, imputed', ylab='Frequency')

```

4b - Calculate and report the mean and median total number of steps taken per day.

```{r}
meanTotalStepsByDayImputed <- mean(totalStepsByDayImputed)
medianTotalStepsByDayImputed <- median(totalStepsByDayImputed)
print(meanTotalStepsByDayImputed)
print(medianTotalStepsByDayImputed)
```

## Are there differences in activity patterns between weekdays and weekends?

1 - Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating
whether a given date is a weekday or weekend day.


```{r}

activityDataCsvImputed$dateType <-  ifelse(as.POSIXlt(activityDataCsvImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
str(activityDataCsvImputed)

```

2 - Make a panel plot containing a time series plot

```{r}
averagedActivityDataCsvImputed <- aggregate(steps ~ interval + dateType, data=activityDataCsvImputed, mean)
str(averagedActivityDataCsvImputed)

ggplot(averagedActivityDataCsvImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("Average number of steps")
```